\Huge \bf \sc Implementation of a quadrotor UAV

### 2012-07-16 latex/paper/paper.tex
vanzo un poquito con la parte del modelo, voy a hacer mear al perro.

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{\Huge \bf
Implementation of a quadrotor UAV}

@@

\begin{abstract}
%\boldmath
This paper describes the design and integration of a control system that allows the autonomous flight of a radio controlled commercial quadcopter. A mathematical model for the quadcopter is developed, its parameters determined from the characterization of the unit. An external intelligence is integrated to play the rol of the flight controller. A 9 degrees of freedom Inertial Measurement Unit (IMU) equipped with a barometer is calibrated and added to the platform. Data from the IMU is combined with the information provided by a GPS within a Extended Kalman Filter (EKF) to obtain a reliable estimation of the state variables. The control actions are obtained from a proportional-integral controller based on the LQR algorithm. For windless environments, a stable platform is achieved.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

A quadcopter platform can be imagined performing countless tasks. Many applications are being developed based on such platforms, but how do they fly? A aerial vehicle is inherently unstable, staying still is not a simple task. This paper focuses on the stabilization of a quadcopter, explaining how to work with an IMU\cite{bib:calib_imu}\cite{bib:bola}\cite{bib:alain}, the development of the mathematical model \cite{bib:quadrotor-bible}\cite{bib:curso_quad} that represents the system, the filtering techniques applied to sensor data\cite{bib:kalman}\cite{bib:kalman2}, and the control system\cite{bib:quadrotor-bible}\cite{bib:hakas} that will allows to quadcopter to fly.

The platform is based on a commercial quadcopter, the frame and the motor control system were preserved, but the IMU and intelligence were replaced by the flight controller that was developed. A BeagleBoard running Linux performs the computations requiered to convert raw data received from the IMU and combine it using an EKF, overcoming the problems inherent to each sensor and the noise generated by the vibrations of the motors, and providing a reliable estimation of the state vector. Once the current state is known, the LQR algorithm\cite{bib:lqrnotes}\cite{bib:lqr-discreto} is used to derive control actions required to bring the system to the desired setpoint.

A diagram of the quadcopter is shown in figure \ref{fig:quad}. Two of the motors rotate clockwise (2 and 4) and the other two (1 and 3) rotate clockwise. This configuration allows the quadcopter to rotate, tilt and gain/lose altitud.

\begin{figure}
	\centering
	\includegraphics[width=.7\columnwidth]{./pics_paper/quad_coord.jpg}
	\caption{\textbf{Model of the quadcopter -} Intertial reference frame in blue, non-inertial reference frame in red. Cyan looped arrows indicate the direction of rotation of each motor. The semicircle and the two yellow spheres indicate the front of the unit.}
%        \vspace{-30pt}
	\label{fig:quad}
\end{figure}

@@

The equations that govern the dynamics of a quadcopter are presented in equation \ref{eq:modelo}. The reference frames used are shown in figure \ref{fig:quad}.

\begin{equation}
\boxed{\begin{aligned}&\dot{x}=v_{q_x} \cos \varphi \cos \theta + v_{q_y} ( \cos \theta \sin \varphi \sin \psi-\cos \varphi \sin \theta )\\
&\quad + v_{q_z}(\sin \psi \sin \theta + \cos \psi \cos \theta \sin \varphi)\\
&\dot{y}=v_{q_x} \cos \varphi \sin \theta + v_{q_y} (\cos \psi \cos \theta + \sin \theta \sin \varphi \sin \psi)\\
&\quad+ v_{q_z}( \cos \psi \sin \theta \sin \varphi-\cos \theta \sin \psi )\\
&\dot{z}= -v_{q_x} \sin \varphi  + v_{q_y} \cos \varphi \sin \psi  + v_{q_z}\cos \varphi \cos \psi\\
&\dot{\psi}=\omega_{q_x} + \omega_{q_z}\tan\varphi \cos\psi + \omega_{q_y}\tan\varphi \sin\psi\\
&\dot{\varphi}=\omega_{q_y}\cos \psi - \omega_{q_z}\sin\psi\\
&\dot{\theta}=\omega_{q_z} \frac{\cos\psi}{\cos\varphi}  + \omega_{q_y}\frac{\sin\psi}{\cos\varphi}\\
&\dot{v_{q_x}}=v_{q_y} \omega_{q_z} - v_{q_z} \omega_{q_y}+g\sin\varphi\\
&\dot{v_{q_y}}=v_{q_z} \omega_{q_x} - v_{q_x} \omega_{q_z}-g\cos\varphi\sin\psi\\
&\dot{v_{q_z}}=v_{q_x} \omega_{q_y} - v_{q_y} \omega_{q_x}-g\cos\varphi\cos\psi+\frac{1}{M}\sum_{i=1}^4T_i\\
&\dot{\omega_{q_x}}=\frac{1}{I_{xx}}\omega_{q_y}\omega_{q_z}(I_{yy}-I_{zz})+\frac{1}{I_{xx}}\omega_{q_y}I_{zz_m}(\omega_1-\omega_2+\omega_3-\omega_4)\\
&\quad-\frac{1}{I_{xx}}dMg\cos\varphi\sin\psi+\frac{1}{I_{xx}}L(T_2-T_4)\\
&\dot{\omega_{q_y}}=\frac{1}{I_{yy}}\omega_{q_x}\omega_{q_z}(-I_{xx}+I_{zz})+\frac{1}{I_{yy}}\omega_{q_x}I_{zz_m}(\omega_1-\omega_2+\omega_3-\omega_4)\\
&\quad-\frac{1}{I_{yy}}dMg\sin\varphi+\frac{1}{I_{yy}}L(T_3-T_1)\\
&\dot{\omega_{q_z}}=\frac{1}{I_{zz}}\left(-Q_1+Q_2-Q_3+Q_4\right)\\
%\frac{-I_{zz_m}(\dot{w_1}-\dot{w_2}+\dot{w_3}-\dot{w_4})}{I_{zz}}+
\end{aligned}}
\label{eq:modelo}
\end{equation}

@@

Two different types of data will be available, depending on the availability of GPS information. When no GPS data is available there will be no direct feedback from the sensors to correct the position on the horizontal plane, only through integration, so only the prediction step can be performed for these variables. As shown in figure \ref{fig:xyz.pdf}, the estimation will drift until GPS data is available.

>>>>

\usepackage{subfig}
%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

@@

% lindo
\newcommand{\refp}[1]{(\ref{#1})}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{\Huge \bf \sc
Implementation of a quadrotor UAV}

@@

\begin{abstract}
%\boldmath
This paper describes the design and integration of a control system that allows the autonomous flight of a commercial quadcopter. A mathematical model for the quadcopter is developed, its parameters determined from the characterization of the unit. An external intelligence is integrated to play the rol of the flight controller. A 9 degrees of freedom Inertial Measurement Unit (IMU) equipped with a barometer is calibrated and added to the platform. Data from the IMU is combined with the information provided by a GPS within a Extended Kalman Filter (EKF) to obtain a reliable estimation of the state variables. The control actions are obtained from a proportional-integral controller based on the LQR algorithm. For windless environments, a stable platform is achieved.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

A quadcopter platform can be imagined performing countless tasks. Many applications are being developed based on such platforms, but how do they fly? A aerial vehicle is inherently unstable, staying still is not a simple task. This paper focuses on the stabilization of a quadcopter, explaining how to work with an IMU\cite{bib:calib_imu}\cite{bib:bola}\cite{bib:alain}, the development of the mathematical model \cite{bib:quadrotor-bible}\cite{bib:curso_quad} that represents the system, the filtering techniques applied to sensor data\cite{bib:kalman}\cite{bib:kalman2}, and the control system\cite{bib:quadrotor-bible}\cite{bib:hakas} that allows the quadcopter to fly.

The platform is based on a commercial radio controlled quadcopter, the frame and the motor control system were preserved, the IMU and intelligence were replaced by the flight controller that was developed. A BeagleBoard\footnote{\url{http://beagleboard.org/}} running Linux\footnote{Angstrom distribution: \url{http://www.angstrom-distribution.org/}} performs the computations requiered to convert raw data received from the IMU over a UART and combine it using an EKF, overcoming the problems inherent to each sensor and the noise generated by the vibrations of the motors, and providing a reliable estimation of the state vector. Once the current state is known, the LQR algorithm\cite{bib:lqrnotes}\cite{bib:lqr-discreto} is used to derive the control actions required to bring the system to the desired setpoint.

\begin{figure}
	\centering
	\includegraphics[width=.7\columnwidth]{./pics_paper/quad_coord.jpg}
	\caption{\textbf{Model of the quadcopter -} In blue arrows represent the intertial reference frame $S_I - \lbrace \hat{i},\hat{j},\hat{k}\rbrace$ ($\lbrace \vec{x},\vec{y},\vec{z}\rbrace$), mapped to  North, West and Up respectively. The red arrows represent the non-inertial reference frame $S_q - \lbrace \hat{i}_q,\hat{j}_q,\hat{k}_q\rbrace$ ($\lbrace \vec{x}_q,\vec{y}_q,\vec{z}_q\rbrace$) relative to the quadcopter. Cyan looped arrows indicate the direction of rotation of each motor (right hand rule). The semicircle and the two yellow spheres indicate the front of the unit.}
%        \vspace{-30pt}
	\label{fig:quad}
\end{figure}

@@

A diagram of the quadcopter is shown in figure \refp{fig:quad}. Two of the motors rotate clockwise (2 and 4) and the other two (1 and 3) rotate clockwise. This configuration allows the quadcopter to rotate, tilt and gain/lose altitud. Two frames of reference are constantly used throught this work: An intertial frame $S_I$ (relative to the Earth) $\lbrace \hat{i},\hat{j},\hat{k}\rbrace - \lbrace \text{North},\text{West},\text{Up}\rbrace$) and a non-intertial frame $S_q$ relative to the quadcopter. Mapping one frame to the other can be achieved by applying the three rotations show in figure \refp{fig:rotaciones}, $\lbrace \theta, \varphi, \psi\rbrace$ are know as Euler angles.

\begin{figure} [h!]
  \centering
  \subfloat[Rotation 1: Axis $\hat{k}$]{\label{fig:angulos_1}
  		\includegraphics[width=0.3\columnwidth]
  			{../pics_modelo_fisico/angulos_1.png}}
  \subfloat[Rotation 2: Axis $\hat{j}$]{\label{fig:angulos_2} 
  		\includegraphics[width=0.3\columnwidth]
  			{../pics_modelo_fisico/angulos_2.png}}
  \subfloat[Rotation 3: Axis $\hat{i}$]{\label{fig:angulos_3} 
  		\includegraphics[width=0.3\columnwidth]
  			{../pics_modelo_fisico/angulos_3.png}}
  \caption{Rotations applied on $S_I$ to obtain $S_q$.}
  \label{fig:rotaciones}
\end{figure}

From a detailed analysis of the dynamics and kinemtics of the quadcopter, the state vector $\hat{X}$ given by (\refp{eq:sv}) is built to describe the system at any given time. Practical considerations lead to choosing the state variables referenced to the quadrotor frame $S_q$. This choice introduces simplifications introduced in the theoretical development, and simplifies interpretation of the data provided by the IMU, which is mounted on the quadcopter and hence provides provides accelerations and angular velocities relative to $S_q$:

\begin{equation}
\hat{X}=\left\lbrace  x,y,z, \theta,\varphi,\psi, v_{q_z},v_{q_y},v_{q_z},\omega_{q_x},\omega_{q_y},\omega_{q_z} \right\rbrace
\label{eq:sv}
\end{equation}

where $\lbrace x,y,z \rbrace$ are the cartesian coordinates relative to the inertial frame $S_I$ (mapped on $\lbrace \hat{i},\hat{j},\hat{k}\rbrace$), $\lbrace\theta,\varphi,\psi\rbrace$ are the Euler angles show in (\refp{fig:rotaciones}), $\lbrace v_{q_x},v_{q_y},v_{q_z}\rbrace$ are the linear velocities relative to $S_I$ and $\lbrace w_{q_x},w_{q_y},w_{q_z}\rbrace$ are the angular velocities relative to $S_q$ (right hand rule applied on $\lbrace \hat{i}_q,\hat{j}_q,\hat{k}_q\rbrace$).

The equations that govern the dynamics of a quadcopter are presented in equation \refp{eq:modelo}. The reference frames used are shown in figure \refp{fig:quad}.

The model for the quadcopter described by equations \refp{eq:modelo} is highly non-linear

\begin{equation}
\small
\boxed{\begin{aligned}&\dot{x}\!=\!v_{q_x} \cos \varphi \cos \theta \!+\! v_{q_y} ( \cos \theta \sin \varphi \sin \psi\!-\!\cos \varphi \sin \theta )\\
&\quad \!+\! v_{q_z}(\sin \psi \sin \theta \!+\! \cos \psi \cos \theta \sin \varphi)\\
&\dot{y}\!=\!v_{q_x} \cos \varphi \sin \theta \!+\! v_{q_y} (\cos \psi \cos \theta \!+\! \sin \theta \sin \varphi \sin \psi)\\
&\quad\!+\! v_{q_z}( \cos \psi \sin \theta \sin \varphi\!-\!\cos \theta \sin \psi )\\
&\dot{z}\!=\! \!-\!v_{q_x} \sin \varphi  \!+\! v_{q_y} \cos \varphi \sin \psi  \!+\! v_{q_z}\cos \varphi \cos \psi\\
&\dot{\psi}\!=\!\omega_{q_x} \!+\! \omega_{q_z}\tan\varphi \cos\psi \!+\! \omega_{q_y}\tan\varphi \sin\psi\\
&\dot{\varphi}\!=\!\omega_{q_y}\cos \psi \!-\! \omega_{q_z}\sin\psi\\
&\dot{\theta}\!=\!\omega_{q_z} \frac{\cos\psi}{\cos\varphi}  \!+\! \omega_{q_y}\frac{\sin\psi}{\cos\varphi}\\
&\dot{v_{q_x}}\!=\!v_{q_y} \omega_{q_z} \!-\! v_{q_z} \omega_{q_y}\!+\!g\sin\varphi\\
&\dot{v_{q_y}}\!=\!v_{q_z} \omega_{q_x} \!-\! v_{q_x} \omega_{q_z}\!-\!g\cos\varphi\sin\psi\\
&\dot{v_{q_z}}\!=\!v_{q_x} \omega_{q_y} \!-\! v_{q_y} \omega_{q_x}\!-\!g\cos\varphi\cos\psi\!+\!\frac{1}{M}\sum_{i\!=\!1}^4T_i\\
&\dot{\omega_{q_x}}\!=\!\frac{1}{I_{xx}}\omega_{q_y}\omega_{q_z}(I_{yy}\!-\!I_{zz})\\
&\quad\!+\!\frac{1}{I_{xx}}\omega_{q_y}I_{zz_m}(\omega_1\!-\!\omega_2\!+\!\omega_3\!-\!\omega_4)\\
&\quad\!-\!\frac{1}{I_{xx}}dMg\cos\varphi\sin\psi\!+\!\frac{1}{I_{xx}}L(T_2\!-\!T_4)\\
&\dot{\omega_{q_y}}\!=\!\frac{1}{I_{yy}}\omega_{q_x}\omega_{q_z}(\!-\!I_{xx}\!+\!I_{zz})\\
&\quad\!+\!\frac{1}{I_{yy}}\omega_{q_x}I_{zz_m}(\omega_1\!-\!\omega_2\!+\!\omega_3\!-\!\omega_4)\\
&\quad\!-\!\frac{1}{I_{yy}}dMg\sin\varphi\!+\!\frac{1}{I_{yy}}L(T_3\!-\!T_1)\\
&\dot{\omega_{q_z}}\!=\!\frac{1}{I_{zz}}\left(\!-\!Q_1\!+\!Q_2\!-\!Q_3\!+\!Q_4\right)\\
%\frac{\!-\!I_{zz_m}(\dot{w_1}\!-\!\dot{w_2}\!+\!\dot{w_3}\!-\!\dot{w_4})}{I_{zz}}\!+\!
\end{aligned}}
\label{eq:modelo}
\end{equation}

@@

Two different types of data will be available, depending on the availability of GPS information. When no GPS data is available there will be no direct feedback from the sensors to correct the position on the horizontal plane, only through integration, so only the prediction step can be performed for these variables. As shown in figure \refp{fig:xyz.pdf}, the estimation will drift until GPS data is available.
